\documentclass{article}


% if you need to pass options to natbib, use, e.g.:
    % \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2022


% ready for submission
\usepackage[preprint]{neurips_2025}


% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2022}


% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2022}


% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2022}


\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\newcommand{\new}[1]{{\color{red} #1}}

\title{Final Project Report For ECE285}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.


\author{%
  Group Number \#: First Name, Last Name
  % examples of more authors
  \And
  First Name, Last Name \\
  \And
  First Name, Last Name \\
  \And
  First Name, Last Name \\
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}




\begin{document}

\maketitle

\begin{abstract}
Please write your abstract here. It should be a snapshot of the entire report.
\end{abstract}


\section{Introduction}
Describe the problem setup and motivation.
Summarize the main research question and contributions.

\section{Related Work}
Review the most relevant prior work.
Position your approach relative to existing methods.

\section{Methodology}
Describe the proposed approach in detail, including problem formulation, model or algorithm design, and implementation considerations.

\section{Experiments}
Describe experimental setup (datasets, baselines, evaluation metrics, etc).

\subsection{Main Results}
Present the primary experimental results using tables and figures.

\subsection{Additional Analysis}
Include ablation studies, sensitivity analysis, or qualitative results as appropriate.

\subsection{Discussion}
Interpret the results.
Discuss what worked well and what did not.

\section{Limitations and Future Work}
Discuss limitations of the current approach and possible directions for improvement.

\section{Conclusion}
Summarize key findings and takeaways.


\appendix
\section{Team Roles and Individual Contributions}
This section outlines the roles of each team member, summarizes their specific contributions, and explains whether the distribution of work was fair, with supporting reasons.


\section{Additional Results or Implementation Details}
Optional.


\bibliography{ref}
\bibliographystyle{abbrvnat}








\end{document}
